---
title: "Homework #2"
author: Elisabeth Gangwer
format:
  html:
    toc: true
    toc-location: left
    theme: custom.scss
    highlight-style: oblivion
    self-contained: true
editor: 
  markdown: 
    wrap: 72
---

## The Data

There is 1000 observations, spread across 10 csv files in a zip.

1.   First you will need to use `unzip` to unzip it.

2.   Next, you will need to generate a vector of all of those files with
    `list.files`. It works like this:

```{r}
library(stringr)
library(reshape2)
library(ggplot2)
file_list <- list.files("~/Desktop/DataWrangling/homework2/data/", pattern = ".csv", 
                        all.files = TRUE, full.names = TRUE)
df_list <- lapply(file_list, read.csv)

better_df <- do.call(rbind, df_list)
```

3.  You can use a `for loop`, an `lapply`, or a `purrr::map` statement
    to bring all of those files together into one data frame. The
    columns are in the same order, so feel free to bind them together.
    If you end up with a list of data frames, you can use
    `do.call(rbind, your_object)`, `dplyr::bind_rows`, or
    `data.table::rbindlist` to bring them together.

## Data Prep Tasks

This part should live in a separate qmd file.

1.  Create a function to clean ratings & length variables.

2.  Create a total length (as in movie length) column.

3.  Create a date delta between release time and review time.

4.  Pick a word from the reviews and count how many times that word
    appears.

5.  Create an aggregated data set for the unique movies. There should be
    movies, average ratings, release year, genre, and total number of
    reviews.

6.  You should have two data frames, so `save` those objects for the
    next step.

## Analytic Tasks

The tasks below should live in a separate qmd from your prep work. You
will want to `load` your data from the previous step.

```{r}
# load Data Prep
load("homework_2_data.Rdata")
# better_df
# even_better_df
```

1.  Which genre has the highest critic rating? Viewer rating?

```{r}
# Create data frame based on genre & critic rating & viewer rating from better_df
agg_rating_maybe <- aggregate(avg_rating ~ reviewer_type + genre, 
                              FUN = mean, 
                              data = better_df)

# Go from long data frame to wide data frame, use dcast to widen data frame  
widen_agg_rating <- dcast(data = agg_rating_maybe, 
                          formula = genre ~ reviewer_type,
                          value.var = "avg_rating")

# Find the highest rating of each reviewer type

# Comedy has the highest average critic rating 
widen_agg_rating[which.max(widen_agg_rating$Critic), ]

# Drama has the highest average viewer rating 
widen_agg_rating[which.max(widen_agg_rating$Viewer), ] 
```

The genre with the highest critic rating is Comedy, while Drama has the
highest viewer rating.

2.  What is the relationship between movie length and average rating?

```{r}
# Movie length is different for movies in the data provided, 
# decided to average the total length in minutes
better_df[better_df$title == 'A Glimpse Through Time', ]


# Create separate data frames based on movie length & average rating, then merge together
# aggregate
agg_movie_rating <- aggregate(avg_rating ~ title, 
                                     FUN = mean, 
                                     data = better_df)
agg_movie_length <- aggregate(total_length ~ title, 
                                     FUN = mean, 
                                     data = better_df)
# merge two aggregations together
movie_rating_length <- merge(x = agg_movie_length,
                              y = agg_movie_rating, 
                              by = 'title')

# Find the correlation between movie length & avg_rating
cor(movie_rating_length$avg_rating, movie_rating_length$total_length)
# Correlation = -0.04958 
# Very low correlation

# plot the movie length & avg_rating to visualize relationship 
ggplot(movie_rating_length, aes(avg_rating, total_length)) +
  geom_point() +
  geom_smooth(method = 'lm', se = FALSE) +
  theme_minimal()
```

The relationship between the total length of a movie and the average
rating is weak. This shows that the total length of a movie does not
necessarily contribute to the average rating.

3.  What is the relationship between the date delta and average rating?

```{r}
# Aggregate date_diff by title from better_df
date_agg_relationship <- aggregate(date_diff ~ title,
                                      FUN = mean, 
                                      data = better_df)

# merge aggregation & avg_rating from even_better_df
rating_date_dif_relationship <- merge(x = date_agg_relationship, 
                                      y = even_better_df,
                                      by = 'title')

# Checking to make sure mean of date diff was calculated correctly 
mean(better_df$date_diff[better_df$title == 'A Glimpse Through Time'])

# Find correlation of avg_rating & date_diff
cor(rating_date_dif_relationship$avg_rating, rating_date_dif_relationship$date_diff)
# correlation = -0.242

# ggplot relationship 
ggplot(rating_date_dif_relationship, aes(avg_rating, date_diff)) +
  geom_point() +
  geom_smooth(method = 'lm', se = FALSE) +
  theme_minimal()
```

The correlation between average rating and date delta is higher than the
length of a movie and the average rating but is still overall a weak
relationship. This suggests that the average rating is not affected by
the date difference of review date and the movie release date.

4.  What is the relationship between total number of reviews and
    average?

```{r}
str(even_better_df)
# convert total reviews to numeric 
even_better_df$total_reviews <- as.numeric(even_better_df$total_reviews)

# Find correlation of total number of reviews & average rating 
cor(even_better_df$avg_rating, even_better_df$total_reviews)
# correlation = 0.475

# Use ggplot to visualize the relationship 
ggplot(even_better_df, aes(avg_rating, total_reviews)) + 
  geom_point() +
  geom_smooth(method = 'lm', se = FALSE) +
  theme_minimal()
```

The relationship between the total number of reviews and average rating
has a positive correlation of 0.475. This suggests, that the more
reviews a movie has the average rating is more likely to increase.

4.  Which movie contains your word of interest at the highest frequency.
    Does that word have any relationship with average rating?

```{r}
# aggregate disaster by title.
# To add the count of disaster not the mean, use the sum function
agg_disaster <- aggregate(disaster ~ title, 
                          data = better_df, 
                          FUN = sum)

# merge aggregation for avg_rating & disaster count by title 
merge_disaster <- merge(x = agg_disaster, 
                      y = even_better_df,
                      by = 'title')

# sort/order the count of disaster 
merge_disaster <- merge_disaster[order(merge_disaster$disaster, decreasing = TRUE), ]
merge_disaster[1:2, ]
# Beyond the Horizon and Sweet Serenade both have the highest frequency of 'disaster' in the review text. 

# Find relationship between avg_rating & 'disaster'
cor(merge_disaster$avg_rating, merge_disaster$disaster)
# correlation is -0.3997

# Make a linear model of avg_rating and disaster to understand relationship. 
lm1 <- lm(avg_rating ~ disaster, data = merge_disaster)
summary(lm1)
```

Although, the amount of times disaster appears in a review would seem
like it would affect the average rating, it does not. I decided to
conduct a linear model. The linear model produced a low r-squared value,
the amount of times disaster appears in movie reviews is insignificant
or does not affect the average rating of a movie.
