---
title: "Homework #1"
format:
  html:
    toc: true
    toc-location: left
    theme: custom.scss
    highlight-style: oblivion
    self-contained: true
editor: 
  markdown: 
    wrap: 72
---
 
See all that stuff above. You should change `theme` and `highlight-style`
to fit your own style. You can find them here:
https://quarto.org/docs/output-formats/html-themes.html

Seriously, you should start finding your own style, so definitely change
them!

## Data

You'll be working with the following data files:

-   `artist_genre.txt`

-   `artist_name.csv`

-   `artist_net_worth.csv`

-   `artist_number_one_hits.csv`

-   `artist_studio_albums.csv`

-   `artist_top_hits.csv`

The only file that you are going to need to be careful about is the
`artist_genre` file -- be sure to look at the extension, refer to the
notes on I/O, and read the help file!

## Goals

Read the files in and then accomplish the following goals:

### Pre-join Goals

-   Use the aggregate function to ensure that each file only contains
    unique rows.

-   If duplicates exist, handle them. You can either de-duplicate them
    (the `duplicate` function is helpful here) or `aggregate` them to
    single rows. The choice is yours and yours alone, but be sure to
    justify your decision.

-   Which 5 artists have the greatest net worth?

-   Which 15 artists have the most number 1 hits?

-   Which 10 artists have the fewest studio albums?

After you have those done, you can join your data frames together.
<br> **Reading in the Data Tables** <br>
```{r}


artist_name <- read.csv("~/Desktop/DataWrangling/homework1/artist_name.csv")

artist_net_worth <- read.csv("~/Desktop/DataWrangling/homework1/artist_net_worth.csv")

artist_number_one_hits <- read.csv("~/Desktop/DataWrangling/homework1/artist_number_one_hits.csv")

artist_studio_albums <- read.csv("~/Desktop/DataWrangling/homework1/artist_studio_albums.csv")

artist_top_hits <- read.csv("~/Desktop/DataWrangling/homework1/artist_top_hits.csv")

artist_genre <- read.table(file = "~/Desktop/DataWrangling/homework1/artist_genre.txt", header = TRUE)

```

<br> **Aggregating Artist Name, removing the duplicates** <br>

```{r}
artist_name_2 <- artist_name

artist_name_2 <- aggregate(Birth_Year ~ Artist, 
          data = artist_name_2,
          FUN = unique)
duplicated(artist_name_2)
```

<br> **Aggregating Artist Net Worth, removing duplicates** <br>

```{r}
artist_net_worth_2 <- artist_net_worth
artist_net_worth_2 <- aggregate(net_worth_millions ~ Artist, 
                                data = artist_net_worth_2,
                                FUN = unique)
```

<br> **Aggregating Artist Genre, removing duplicates** <br>

```{r}
artist_genre_2 <- artist_genre
artist_genre_2 <- aggregate(genre ~ Artist, 
                            data = artist_genre_2, 
                            FUN = unique)
```

<br> **Aggregating Artist number one hits, removing duplicates** <br>

```{r}
artist_number_one_hits_2 <- artist_number_one_hits
artist_number_one_hits_2 <- aggregate(number_one_hits ~ artist, 
                            data = artist_number_one_hits_2, 
                            FUN = unique)
```

<br> **Aggregating Artist Studio Albums, removing duplicates** <br>

```{r}
artist_studio_albums_2 <- artist_studio_albums
artist_studio_albums_2 <- aggregate(studio_albums ~ Artist, 
                                    data = artist_studio_albums_2, 
                                    FUN = unique)
```

<br> **Aggregating Top Hits by Artist, removing duplicates** <br>

```{r}
artist_top_hits_2 <- artist_top_hits
artist_top_hits_2 <- aggregate(cbind(Top_Hit, Year) ~ Artist, 
                               data = artist_top_hits_2, 
                               FUN = unique)
```

<br> **Sorting Artist by their Net Worth in Millions to determine the
top 5 artists w/ highest net worth** <br>

```{r}
artist_net_worth_2$net_worth_millions <- 
  as.numeric(gsub("\\$", "", x = artist_net_worth_2$net_worth_millions))
sort(artist_net_worth_2$net_worth_millions, decreasing = TRUE)
artist_net_worth_2 <- artist_net_worth_2[order(artist_net_worth_2$net_worth_millions, 
                             decreasing = TRUE),]
artist_net_worth_2[1:5, ]
```

-   Which 5 artists have the greatest net worth?

1.  Jay-Z (\$1500M)
2.  The Beatles (\$1500M)
3.  ABBA (\$900M)
4.  The Rolling Stones (\$900M)
5.  Madonna (\$850M)

<br> **Sorting Artists by their #1 Hits to determine top 15 artists.**
<br>

```{r}
artist_number_one_hits_2 <- artist_number_one_hits_2[order(artist_number_one_hits_2$number_one_hits, 
                             decreasing = TRUE), ]
artist_number_one_hits_2[1:15, ]
```

-   Which 15 artists have the most number 1 hits?

1.  The Beatles (20)
2.  Mariah Carey (19)
3.  Elvis Presley (18).
4.  Rihanna (14)
5.  Michael Jackson(13) )
6.  Madonna (12)
7.  Whitney Houston (11)
8.  Stevie Wonder (10)
9.  Katy Perry (9)
10. Drake (8)
11. The Rolling Stones (8)
12. Bruno Mars(7)
13. Taylor Swift (7)
14. Ed Sheeran (6)
15. Elton John (6)

<br> **Which artists have the fewest studio albums?** <br>

```{r}
artist_studio_albums_2 <- artist_studio_albums_2[order(artist_studio_albums_2$studio_albums,
                             decreasing = FALSE), ]
artist_studio_albums_2[1:10, ]
```

-   Which 10 artists have the fewest studio albums?
    1.  Billie Eillish (1)
    2.  Lorde (2)
    3.  The Notorious B.I.G (2)
    4.  Adele (3)
    5.  Bruno Mars (3)
    6.  Jimi Hendrix (3)
    7.  Nirvana (3)
    8.  Sam Smith (3)
    9.  Imagine Dragons (4)
    10. Justin Timerlake (4)

### Post-join Goals

-   Create variables to capture the following ideas:
    -   How long has the artist been around? You can just use their
        birth year.
    -   Something to capture each artist's net worth per studio album.
-   Plot net worth per studio album, the "been around" variable, and
    genre. There should be points and colors here.

For all tasks comment your code and explain what you found!

<br>**Join Tables** <br>

```{r}
merge_artist_1 <- merge(x = artist_genre_2,
                        y = artist_name_2,
                        by = "Artist",
                        all.x = TRUE)
merge_artist_2 <- merge(x = artist_net_worth_2,
                        y = artist_studio_albums_2,
                        by = "Artist",
                        all.x = TRUE)
merge_artist_3 <- merge(x = merge_artist_1,
                        y= merge_artist_2,
                        by = "Artist",
                        all.x = TRUE)

# Need to change column names of Artist Column in
# Number One Hits and Top Hits to make sense & merge into the complete_artist_data data frame

colnames(artist_number_one_hits_2)[colnames(artist_number_one_hits_2) == "artist"] = "Artist"
colnames(artist_top_hits_2)[colnames(artist_top_hits_2) == "Year"] = "Top_Hit_Release_Year"

merge_artist_4 <- merge(x = artist_top_hits_2,
                        y = artist_number_one_hits_2,
                        by = "Artist",
                        all.x = TRUE)
complete_artist_data <- merge(x = merge_artist_3,
                              y = merge_artist_4,
                              by = "Artist",
                              all.x = TRUE)
```

<br> **Calculate how long the artist has 'been around' or their age
Column Name will be Artist Been Around in Complete Artist Data** <br>

```{r}
library(dplyr)
current_year <- as.numeric(format(Sys.Date(),"%Y")) 
current_year
artist_been_around <- current_year - complete_artist_data$Birth_Year

artist_been_around <- data.frame(artist_name_2$Artist, artist_been_around)

colnames(artist_been_around)[colnames(artist_been_around) == "artist_name_2.Artist"] = "Artist"

complete_artist_data <- merge(x = complete_artist_data,
                              y = artist_been_around, 
                              by = "Artist",
                              all.x = TRUE)

```

<br> **Something to capture each artist's net worth per studio album.**
<br>

```{r}
net_worth_studio_album <- merge_artist_2$net_worth_millions / merge_artist_2$studio_albums
net_worth_studio_album <- round(net_worth_studio_album, digits = 2)
#Add Net Worth per Studio Album to Complete Artist Data Set
complete_artist_data$net_worth_studio_album = net_worth_studio_album

```

-   Plot net worth per studio album, the "been around" variable, and
    genre. There should be points and colors here.

```{r}
library(ggplot2)

ggplot(complete_artist_data, aes(x = artist_been_around,
           y = net_worth_studio_album,
           color = genre)) +
      geom_point() +
      theme_bw()
```
